The Error Estimate for Computing Electrostatic Interaction in the Inhomogeneous and Correlated Molecular Systems: for the Ewald Summation, SPME and Staggered Mesh Ewald.

### 2012-04-25 longRange.hetero/publish/main.tex
fix bugs

\documentclass[aps,pre,preprint,unsortedaddress]{revtex4}
% \documentclass[aps,pre,unsortedaddress]{revtex4-1}
%\documentclass[aps,pre,twocolumn]{revtex4-1}
%\documentclass[aps,jcp,groupedaddress,twocolumn,unsortedaddress]{revtex4}

@@

\begin{abstract}

\end{abstract}

@@

The calculation of the long-range electrostatic interaction is a very
important topic in the molecular simulations.  Due to the very slow
decay of the interaction with respect to the distance between charges,
the short-range fast algorithms (the cell list and neighbor list
algorithms) do not work, and the
energy of the system is only conditionally convergent.  One of the
first algorithms handling this problem was the Ewald summation, dating
back to the 1920's~\cite{ewald1921die}.  However, it has been shown that
with a careful tuning of working parameters, the optimal computational
cost of the Ewald summation is of $\mathcal
O(N^{3/2})$~\cite{perram1988asc}, which is not feasible for modern
large scale molecular simulations.
Therefore, several Ewald based fast algorithms were proposed to reduce the
computational cost to a scalable level: $\mathcal O(N\log N)$. Some of them
are
the particle mesh Ewald (PME)
method~\cite{darden1993pme}, the smooth particle mesh Ewald (SPME)
method~\cite{essmann1995spm} and the particle-particle-particle-mesh
(P3M) method~\cite{hockney1988computer, deserno1998mue1}.
More recently, the staggered mesh or
``interlacing'' technique~\cite{chen1974reduction,eastwood1976optimal}
was applied to SPME
(called staggered mesh Ewald)~\cite{cerutti2009staggered} and P3M
(called interlaced P3M)~\cite{neelov2010interlaced}, and was showed to
improve the accuracy of these algorithms.
In this paper, we focus on the Ewald summation, the SPME and
the staggered mesh Ewald methods. The Ewald summation is the starting
point of all the mentioned fast algorithm. SPME
is the one of the most popular long-range algorithms among
the mainstream molecular simulation packages, for example,
AMBER~\cite{case2005amber}, GROMACS~\cite{van2005gromacs,
  hess2008gromacs} and NAMD~\cite{phillips2005scalable}.

The aforementioned fast algorithms have achieved great success
over the past three decades.
However, the fact that the user should provide six working parameters
for these methods may be problematic: it is difficult to try out the
most efficient parameter set in such a high dimensional parameter space.
The experiential choices may be either less accurate or waste
computational effort.
This problem can be solved by an
in-depth study of the error introduced by these algorithms, namely the
error estimate~\cite{kolafa1992cutoff, hummer1995numerical,
  petersen1995accuracy, deserno1998mue2, stern2008mesh,
  wang2010optimizing, neelov2010interlaced, ballenegger2012convert},
by which the error is described by a function of the working
parameters.  Using the error estimates, a work flow has been designed to
automatically determine the nearly optimal combination of parameters,
in the sense that it is the fastest one given a demanding
accuracy~\cite{wang2010optimizing}. However, so far, all error
estimates assume the homogeneity of the charge distribution, and the
independency of any pair of charges (i.e. the charges are
uncorrelated), but neither of the assumptions is satisfactorily
fulfilled in most molecular
systems of practical interest.  For example, the lipid biomembrane
solving in water: the carbon chains of the lipids carry no charge, and
the charges are correlated by the covalent bonds, the van der Waals
interaction, the hydrogen bonding network, etc..  Therefore, the current
error estimates may be problematic, or even not applicable.  A lot of
research have shown that due to the correlation of charges, the error estimates
obviously deviate from the real
error~\cite{deserno1998mue2,wang2010optimizing,ballenegger2012convert}.

The purpose of this research is to develop a reliable  error
estimate in inhomogeneous and correlated molecular systems.
We estimate the force error because the application is the molecular
dynamics (MD) simulation.  Firstly, a general error estimate
frame work for the pairwise interaction is set up. We prove that the
error is decomposed into three additive parts: the homogeneity error,
the inhomogeneity error and the correlation error~\cite{wang2012}.
These error estimates are only of theoretical interest, and
not applicable, because the computational costs of the error estimates
are $\mathcal O(N^2\log N)$ (homogeneity and inhomogeneity error)
and $\mathcal O(N^3\log N)$ (correlation error).
If the error force has a special form so that the error estimates
can be written as convolutions, the computational costs
are reduced to $\mathcal O(N\log N)$ and 
$\mathcal O(N^2\log N)$, respectively.
The computational cost of the correlation error is further
reduce to $\mathcal O(N\log N)$ by the nearest neighbor approximation
technique, which assume the system is only short-range correlated. This
is the case for most the liquid and gas system far from critical point.
Following this framework, we
develop the error estimates for the Ewald summation, the SPME method
and the staggered method Ewald method.   The estimates
of the homogeneity and inhomogeneous errors are tested by two
randomly distributed inhomogeneous charges systems
(i.e. the charges are not correlated), and is proved to
be very sharp.
The error estimates are also tested in a real inhomogeneous water system,
in which the charge correlation plays a very important role
on the force error. By 
including the nearest neighbor approximation of
correlation error, the quality of the estimates
is impressively enhanced.
This paper is ended by conclusions and remarks in
section~\ref{sec:conclusions}.

@@

In this paper, we consider the system with periodic boundary condition.
Suppose it is
composed by $N$ charged particles located at $\v r_1, \v r_2, \cdots,
\v r_N$ with charges $q_1, q_2, \cdots, q_N$, respectively.
We study the force of a testing particle located at $\v r$, which
feels interactions from all charges $q_1, q_2, \cdots, q_N$
in the system, and does not
exert force on them.
The \emph{error force} is defined by the difference between
the exact force and the calculated
force on the testing particle
\cite{wang2012}, and is denoted by $\Delta \v F(\v r)$.
The magnitude of the force computation error 
is defined by the \emph{root mean square (RMS)} \emph{error}, which is
the square root of the second moment of the error force: $\mathcal
E(\v r) = \sqrt{\langle\vert\Delta\v F(\v r)\vert^2\rangle}$.  The
notation $\langle\cdot\rangle$ means ensemble average.  We also
estimate the the ensemble average of the error force (also called
\emph{mean error force}), namely $\langle\Delta\v F(\v r)\rangle$,
which plays an important role in the error analysis of the
inhomogeneous systems~\cite{wang2012}.

@@

    \mathcal E (\v r) 

@@

  systems~\cite{deserno1998mue2, wang2010optimizing}.

@@

  By setting $q_i = q_j = 1$, the corollary~\ref{thm:tmp2} directly proves

@@

@@

  Without the property $\v K(\v r,\v r') = \v K(\v r - \v r')$,
  theorem~\ref{thm:tmp1} calculates the homogeneity and inhomogeneity

@@

  in inhomogeneity systems~\cite{wang2012}.

@@

The full computation of the correlation error cost at least $\mathcal

@@

\textrm{atom $k$ is the nearest neighbor of atom $j$} \,\}$,

@@

The complementary error function $\textrm{erfc}(r)$ in
Eqn.~\eqref{eqn:tmp2} decays exponentially as $r$ goes to infinity.
Therefore, it is a short-range interaction in the real space, which
can be calculated by the standard cut-off and neighbor list
method~\cite{frenkel02b} at a cost of $\mathcal O(N)$.  The reciprocal
part also decays exponentially as the magnitude of the Fourier mode
$\vert\v m\vert$ increases. Therefore, in practice, the infinite
summation in Eqn.~\eqref{eqn:tmp3} is truncated, and only a finite
summation is calculated:
% Both the direct and the reciprocal parts are calculated by cut-off in
% real and reciprocal space, respectively.
% For the reciprocal part the cut-offed energy reads
\begin{align}\label{eqn:tmp6}
  E^{\textrm{tr}}_{\textrm{rec}} & =
  % \frac1{2\pi V}
  \sum_{
    \begin{subarray}{c}
      \vert m_\alpha\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  % \frac{\exp(-\pi^2\v m^2 / \beta^2)}{\v m^2} S(\v m) S(-\v m)   
  f(\v m)\, S(\v m) S(-\v m),
\end{align}
where, for short,
\begin{align}
  f(\v m) := \frac{\exp(-\pi^2 \v m^2 / \beta^2)}{2\pi V \v m^2},
\end{align}
and $K_\alpha$ is the number of Fourier modes used on direction
$\alpha$.  The truncated reciprocal force acting on particle $i$ is
\begin{align}\nonumber
  \v F^{\textrm{tr}}_{\textrm{rec}}(\v r_i)
  &= 
  q_i 
  \sum_{
    \begin{subarray}{c}
      \vert m_\alpha\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  \v g(\v m) \,
  S(-\v m)\,
  e^{2\pi i\v m\cdot\v r_i}, \\\label{eqn:tmp8}
  &= 
  \sum_j   q_iq_j
  \sum_{
    \begin{subarray}{c}
      \vert m_\alpha\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  \v g(\v m) \,
  e^{2\pi i\v m\cdot\v r_{ij}},
\end{align}
where
\begin{align}
  \v g(\v m) = -4 \pi \v m i f(\v m).
\end{align}

@@

The error force of the ik-differentiation is therefore
  \sum_{j}q_j\,e^{-2\pi i\v m\cdot\v r_j} \\\nonumber
  \sum_{j}q_j\,A(\v m, -\v r_j)\,e^{-2\pi i\v m\cdot\v r_j}\\\label{eqn:ik.error.force.1}
  A(\v m, \v r) +
  A(\v m, -\v r)
  \big].
Due to the symmetricity of $\v g(\v m)$ and $A(\v m, \v r)$,
the last term in Eqn.~\eqref{eqn:ik.error.force.1},
i.e. the ``self-interacting term'', is zero.
Therefore,

@@

error force kernel of ik-differentiation

@@

possible to develop the error estimate with convolution form.

@@

The nearest neighbor approximation to the correlation error is:

@@

The error force of the analytical differentiation is 
\begin{align}\nonumber
  \Delta\v F_{\textrm{rec}}^{\textrm{ana}}(\v r)
  % =\;&
  % q
  % \sum_{
  %   \begin{subarray}{c}
  %     \vert m_{\alpha}\vert < K_\alpha/2\\
  %     \v m\neq 0
  %   \end{subarray}}
  % -2f(\v m)\,
  % \nabla_{\v r}
  % e^{2\pi i\v m\cdot \v r}
  % \sum_{j\neq i}q_j\,
  % A(\v m, -\v r_j)
  % e^{-2\pi i\v m\cdot\v r_j} \\\nonumber
  % & +\,
  % q
  % \sum_{
  %   \begin{subarray}{c}
  %     \vert m_{\alpha}\vert < K_\alpha/2\\
  %     \v m\neq 0
  %   \end{subarray}}
  % -2f(\v m)\,
  % \nabla_{\v r}
  % [A(\v m, \v r)\,
  % e^{2\pi i\v m\cdot \v r} ]
  % \sum_{j\neq i}q_j\,
  % e^{-2\pi i\v m\cdot\v r_j} \\\nonumber
  % & +\,
  % q^2
  % \sum_{
  %   \begin{subarray}{c}
  %     \vert m_{\alpha}\vert < K_\alpha/2\\\nonumber
  %     \v m\neq 0
  %   \end{subarray}}
  % -2f(\v m)\,
  % A(-\v m, \v r)\,e^{2\pi i\v m\cdot \v r} \nabla_{\v r}e^{2\pi i\v m\cdot\v r} \\\nonumber
  % & +\,
  % q^2
  % \sum_{
  %   \begin{subarray}{c}
  %     \vert m_{\alpha}\vert < K_\alpha/2\\
  %     \v m\neq 0
  %   \end{subarray}}
  % -2f(\v m)\,
  % e^{-2\pi i\v m\cdot \v r}
  % \nabla_{\v r} [\,A(\v m, \v r) \,e^{2\pi i\v m\cdot\v r}\,]  \\\nonumber
  =&\,
  \Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r)\\\nonumber
  & +\,
  q
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  -2f(\v m)\,
  \v B(\v m, \v r)\,e^{2\pi i\v m\cdot \v r}
  \sum_{j}q_j\,
  e^{-2\pi i\v m\cdot\v r_j} \\\label{eqn:ana.error.force}
  & +\,
  q^2
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  -2f(\v m)\,\v B(\v m, \v r) 
\end{align}
With $\v B(\v m, \v r)$ defined by
\begin{align}\nonumber
  \v B(\v m, \v r)
  &=
  \nabla_{\v r}\,A(\v m, \v r)\\\label{eqn:error-B}
  &=
  \sum_{\alpha=1}^3\sum_{l\neq 0}
  Z_{\alpha,l}(\v m)\,
  2\pi i l K_\alpha\v a_\alpha^\ast\,e^{2\pi i l K_\alpha\v a^\ast_\alpha\cdot \v r} 
\end{align}
The last term in Eqn.~\eqref{eqn:ana.error.force} is due to the
self-interaction~\cite{cerutti2009staggered, ballenegger2011removal,
  neelov2010interlaced}, which does not depends on the coordinates of
particles. Therefore, it can be calculated explicitly, and subtracted
from the analytical force during the simulation. The computational cost of
the self-interaction term is very low comparing with the SPME
force calculation. It is shown that the self-interaction  error
dominate in the low charge density regions~\cite{cerutti2009staggered}.
Therefore, we always remove the self-interaction term from
the analytical differentiation.
% , and call it as the original analytical
% differetiation when comparing it with the staggered mesh analytical
% differetiation.

@@

\subsection{The error estimate of the total force}

>>>>

\documentclass[aps,pre,preprint,unsortedaddress]{revtex4}
% \documentclass[aps,jcp,unsortedaddress]{revtex4-1}
%\documentclass[aps,pre,twocolumn]{revtex4-1}
%\documentclass[aps,jcp,groupedaddress,twocolumn,unsortedaddress]{revtex4}

@@

\begin{abstract}
  In this work, we develop the accurate error estimates for
  some state-of-art algorithms of long-range electrostatic interaction
  in inhomogeneous and correlated molecular systems.
  They are the Ewald summation, the smooth particle mesh Ewald (SPME) and
  the staggered mesh Ewald method. Two branches of force computation, namely
  the ik- and analytical differentiation, are considered. All the
  estimates are developed by proposing a more general framework:
  if the error force is of pairwise form,
  then the root mean square force error is composed by three additive parts,
  the homogeneity error, the inhomogeneity error and the correlation
  error. Computationally scalable estimates (estimating the errors
  at the cost $\mathcal O(N\log N)$) are developed for all the considered
  algorithms.
  The effectiveness of the proposed estimates and the important role of
  the correlation error are carefully checked and demonstrated by
  example systems.
\end{abstract}

@@

The calculation of the long-range electrostatic interaction is a very
important topic in the molecular simulations.  Due to the very slow
decay of the interaction strength with respect to the distance between charges,
the typical short-range fast algorithms (cut-off + neighbor list
algorithms) do not work, and the
energy of the system is only conditionally convergent.  One of the
first algorithms handling this problem is the Ewald summation, dating
back to the 1920's~\cite{ewald1921die}.  It has been shown that
with a careful tuning of working parameters, the optimal computational
cost of the Ewald summation is  $\mathcal
O(N^{3/2})$~\cite{perram1988asc}, which is not feasible for modern
large scale molecular simulations.
Therefore, several Ewald based fast algorithms were proposed to reduce the
computational cost to a scalable level: $\mathcal O(N\log N)$. Some of them
are
the particle mesh Ewald (PME)
method~\cite{darden1993pme}, the smooth particle mesh Ewald (SPME)
method~\cite{essmann1995spm} and the particle-particle-particle-mesh
(P3M) method~\cite{hockney1988computer, deserno1998mue1}.
More recently, the staggered mesh or
``interlacing'' technique~\cite{chen1974reduction,eastwood1976optimal}
was applied to SPME
(called the staggered mesh Ewald)~\cite{cerutti2009staggered} and P3M
(called the interlaced P3M)~\cite{neelov2010interlaced}, and was showed to
improve the accuracy of these algorithms greatly.
In this paper, we focus on the Ewald summation, the SPME and
the staggered mesh Ewald methods. The Ewald summation is the starting
point of all the mentioned fast algorithm. The SPME
is the one of the most popular long-range algorithms among
the mainstream molecular simulation packages, for example,
AMBER~\cite{case2005amber}, GROMACS~\cite{van2005gromacs,
  hess2008gromacs} and NAMD~\cite{phillips2005scalable}.

The aforementioned fast algorithms have achieved great success
over the past three decades.
However, the fact that the user should provide six working parameters
for these methods may be problematic: it is difficult to try out the
most efficient parameter set in such a high dimensional parameter space.
The experiential choices may be either less accurate or waste
computational effort.
This problem can be solved by an
in-depth study of the error introduced by these algorithms, namely the
error estimate~\cite{kolafa1992cutoff, hummer1995numerical,
  petersen1995accuracy, deserno1998mue2, stern2008mesh,
  wang2010optimizing, neelov2010interlaced, ballenegger2012convert},
by which the error is described by a function of the working
parameters.  Using the error estimates, a work flow has been designed to
automatically determine the nearly optimal combination of parameters,
in the sense that it is the fastest one satisfying a demanding
accuracy~\cite{wang2010optimizing}. However, so far, all error
estimates assume the homogeneity of the charge distribution, and the
independency of any pair of charges (i.e. the charges are
uncorrelated), but neither of the assumptions is satisfactorily
fulfilled in most molecular
systems of practical interest.  For example, the lipid bilayer membrane
solving in water: the hydrophobic tail(s) of the lipids carry no charge, and
the charges in the system
are correlated by the covalent bonds, the van der Waals
interaction, the hydrogen bonding network, etc..  Therefore, the current
error estimates may be problematic, or even not applicable.  Several
research have shown that due to the correlation of charges, the error estimates
obviously deviate from the real
error~\cite{deserno1998mue2,wang2010optimizing,ballenegger2012convert}.

The purpose of this research is to develop a reliable  error
estimate in inhomogeneous and correlated molecular systems.
We estimate the force error because the application is the molecular
dynamics (MD) simulation.  Firstly, a general error estimate
framework for the pairwise interaction is set up. We prove that the
error is decomposed into three additive parts: the homogeneity error,
the inhomogeneity error and the correlation error~\cite{wang2012}.
% These error estimates are only of theoretical interest, and
% not applicable, because the computational costs of the error estimates
% are $\mathcal O(N^2)$ (homogeneity and inhomogeneity error)
% and $\mathcal O(N^3)$ (correlation error).
% If the error force has a special form so that the error estimates
% can be written as convolutions, the computational costs
% are reduced to $\mathcal O(N\log N)$ and 
% $\mathcal O(N^2\log N)$, respectively.
The computational cost of the homogeneity and inhomogeneity
error is $\mathcal O(N\log N)$ when the error force kernel is
of convolution form.
The  correlation error can also be
estimated at a cost of $\mathcal O(N\log N)$
by the nearest neighbor approximation
technique, which assume the error  is dominated by  short-range correlation.
This
is the case for most the liquid and gas systems far from critical point.
Generally speaking,
when the error force can be written as a pairwise form, 
all estimate are derived easily under this framework.
Following this framework, we
develop the error estimates for the Ewald summation, the SPME method
and the staggered method Ewald method.   The estimates
of the homogeneity and inhomogeneous errors are tested by two
randomly distributed inhomogeneous charges systems,
% (i.e. the charges are not correlated),
and is proved to
be very sharp.
The error estimates are also tested in a real inhomogeneous water system,
in which the charge correlation plays a very important role
on the force error. By 
including the nearest neighbor approximation of
correlation error, the quality of the estimates
is impressively enhanced.
This paper is ended by conclusions and remarks in
section~\ref{sec:conclusions}.

@@

In this paper, we consider the system with periodic boundary condition.
Suppose it is
composed by $N$ charged particles located at $\v r_1, \v r_2, \cdots,
\v r_N$ with charges $q_1, q_2, \cdots, q_N$, respectively.
We study the force of a testing particle located at $\v r$, which
feels interactions from all charges $q_1, q_2, \cdots, q_N$
in the system, and does not
exert force on them.
The \emph{error force} is defined by the difference between
the exact force and the calculated
force on the testing particle
\cite{wang2012}, and is denoted by $\Delta \v F(\v r)$.
The magnitude of the force computation error 
is defined by the \emph{root mean square (RMS)} \emph{error}, which is
the square root of the second moment of the error force: $\mathcal
E(\v r) = \sqrt{\langle\vert\Delta\v F(\v r)\vert^2\rangle}$.  The
notation $\langle\cdot\rangle$ means the ensemble average.  We also
estimate the the ensemble average of the error force (also called
\emph{mean error force}), namely $\langle\Delta\v F(\v r)\rangle$,
which plays an important role in the error analysis of the
inhomogeneous systems~\cite{wang2012}.

@@

@@

  systems~\cite{deserno1998mue2, wang2010optimizing, ballenegger2012convert}.

@@

  By setting $q_i = q_j = 1$, the corollary~\ref{thm:tmp2} proves

@@

    % \langle\vert\Delta\v F(\v r)\vert^2\rangle
    \mathcal E^2(\v r)

@@

  % Without the property $\v K(\v r,\v r') = \v K(\v r - \v r')$,
  Theorem~\ref{thm:tmp1} calculates the homogeneity and inhomogeneity

@@

  in inhomogeneous systems~\cite{wang2012}.

@@

The full computation of the correlation error costs at least $\mathcal

@@

\textrm{atom $k$ is one of the nearest neighbors of atom $j$} \,\}$,

@@

The complementary error function $\textrm{erfc}(r)$ in
Eqn.~\eqref{eqn:tmp2} decays exponentially as $r$ goes to infinity.
Therefore, it is a short-range interaction in the real space, which
can be calculated by the standard cut-off and neighbor list
method~\cite{frenkel02b} at a cost of $\mathcal O(N)$.  The reciprocal
part also decays exponentially as the magnitude of the Fourier mode
$\vert\v m\vert$ increases. Therefore, in practice, the infinite
summation in Eqn.~\eqref{eqn:tmp3} is truncated, and only a finite
summation is calculated:
% Both the direct and the reciprocal parts are calculated by cut-off in
% real and reciprocal space, respectively.
% For the reciprocal part the cut-offed energy reads
\begin{align}\label{eqn:tmp6}
  E^{\textrm{tr}}_{\textrm{rec}} & =
  % \frac1{2\pi V}
  \sum_{
    \begin{subarray}{c}
      \vert m_\alpha\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  % \frac{\exp(-\pi^2\v m^2 / \beta^2)}{\v m^2} S(\v m) S(-\v m)   
  f(\v m)\, S(\v m) S(-\v m),
\end{align}
where, for short,
\begin{align}
  f(\v m) := \frac{\exp(-\pi^2 \v m^2 / \beta^2)}{2\pi V \v m^2},
\end{align}
and $K_\alpha$ is the number of Fourier modes used on direction
$\alpha$.  The truncated reciprocal force acting on particle $i$ is
\begin{align}
  % \nonumber
  \v F^{\textrm{tr}}_{\textrm{rec}}(\v r_i)
  % &= 
  % q_i 
  % \sum_{
  %   \begin{subarray}{c}
  %     \vert m_\alpha\vert < K_\alpha/2\\
  %     \v m\neq 0
  %   \end{subarray}}
  % \v g(\v m) \,
  % S(-\v m)\,
  % e^{2\pi i\v m\cdot\v r_i}, \\
  \label{eqn:tmp8}
  &= 
  \sum_j   q_iq_j
  \sum_{
    \begin{subarray}{c}
      \vert m_\alpha\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  \v g(\v m) \,
  e^{2\pi i\v m\cdot\v r_{ij}},
\end{align}
where
\begin{align}
  \v g(\v m) = -4 \pi \v m i f(\v m).
\end{align}

@@

Due to symmetricity of $\v g(\v m)$ and $A(\v m, \v r)$, the
``self-interacting'' term of the
ik-differentiation vanishes~\cite{wang2010optimizing}.
The error force is therefore
% \begin{align}\nonumber
%   \Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r)
%   =\;&
%   q
%   \sum_{
%     \begin{subarray}{c}
%       \vert m_{\alpha}\vert < K_\alpha/2\\
%       \v m\neq 0
%     \end{subarray}}
%   \v g(\v m)\,
%   A(\v m, \v r)\,
%   e^{2\pi i\v m\cdot \v r}
%   \sum_{j}q_j\,e^{-2\pi i\v m\cdot\v r_j} \\\nonumber
%   & +\,
%   q
%   \sum_{
%     \begin{subarray}{c}
%       \vert m_{\alpha}\vert < K_\alpha/2\\
%       \v m\neq 0
%     \end{subarray}}
%   \v g(\v m)\,
%   e^{2\pi i\v m\cdot \v r}
%   \sum_{j}q_j\,A(\v m, -\v r_j)\,e^{-2\pi i\v m\cdot\v r_j}\\\label{eqn:ik.error.force.1}
%   & +\,
%   q
%   \sum_{
%     \begin{subarray}{c}
%       \vert m_{\alpha}\vert < K_\alpha/2\\
%       \v m\neq 0
%     \end{subarray}}
%   \v g(\v m)\,
%   \big[
%   A(\v m, \v r) +
%   A(\v m, -\v r)
%   \big].
% Due to the symmetricity of $\v g(\v m)$ and $A(\v m, \v r)$,
% the last term in Eqn.~\eqref{eqn:ik.error.force.1},
% i.e. the ``self-interacting term'', is zero.
% Therefore,

@@

error force kernel

@@

possible to develop the error estimate with convolution form under
some approximations.

@@

The nearest neighbor approximation to the correlation error
in a water system is:

@@

The error force of the analytical differentiation is 
\begin{align}\nonumber
  \Delta\v F_{\textrm{rec}}^{\textrm{ana}}(\v r)
  % =\;&
  % q
  % \sum_{
  %   \begin{subarray}{c}
  %     \vert m_{\alpha}\vert < K_\alpha/2\\
  %     \v m\neq 0
  %   \end{subarray}}
  % -2f(\v m)\,
  % \nabla_{\v r}
  % e^{2\pi i\v m\cdot \v r}
  % \sum_{j\neq i}q_j\,
  % A(\v m, -\v r_j)
  % e^{-2\pi i\v m\cdot\v r_j} \\\nonumber
  % & +\,
  % q
  % \sum_{
  %   \begin{subarray}{c}
  %     \vert m_{\alpha}\vert < K_\alpha/2\\
  %     \v m\neq 0
  %   \end{subarray}}
  % -2f(\v m)\,
  % \nabla_{\v r}
  % [A(\v m, \v r)\,
  % e^{2\pi i\v m\cdot \v r} ]
  % \sum_{j\neq i}q_j\,
  % e^{-2\pi i\v m\cdot\v r_j} \\\nonumber
  % & +\,
  % q^2
  % \sum_{
  %   \begin{subarray}{c}
  %     \vert m_{\alpha}\vert < K_\alpha/2\\\nonumber
  %     \v m\neq 0
  %   \end{subarray}}
  % -2f(\v m)\,
  % A(-\v m, \v r)\,e^{2\pi i\v m\cdot \v r} \nabla_{\v r}e^{2\pi i\v m\cdot\v r} \\\nonumber
  % & +\,
  % q^2
  % \sum_{
  %   \begin{subarray}{c}
  %     \vert m_{\alpha}\vert < K_\alpha/2\\
  %     \v m\neq 0
  %   \end{subarray}}
  % -2f(\v m)\,
  % e^{-2\pi i\v m\cdot \v r}
  % \nabla_{\v r} [\,A(\v m, \v r) \,e^{2\pi i\v m\cdot\v r}\,]  \\\nonumber
  =&\,
  \Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r)\\\nonumber
  & +\,
  q
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  -2f(\v m)\,
  \v B(\v m, \v r)\,e^{2\pi i\v m\cdot \v r}
  \sum_{j}q_j\,
  e^{-2\pi i\v m\cdot\v r_j} \\\label{eqn:ana.error.force}
  & +\,
  q^2
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  -2f(\v m)\,\v B(\v m, \v r) 
\end{align}
With $\v B(\v m, \v r)$ defined by
\begin{align}\nonumber
  \v B(\v m, \v r)
  &=
  \nabla_{\v r}\,A(\v m, \v r)\\\label{eqn:error-B}
  &=
  \sum_{\alpha=1}^3\sum_{l\neq 0}
  Z_{\alpha,l}(\v m)\,
  2\pi i l K_\alpha\v a_\alpha^\ast\,e^{2\pi i l K_\alpha\v a^\ast_\alpha\cdot \v r} 
\end{align}
The last term in Eqn.~\eqref{eqn:ana.error.force} is due to the
self-interaction~\cite{cerutti2009staggered, ballenegger2011removal,
  neelov2010interlaced}, which does not depends on the coordinates of
particles. Therefore, it can be calculated explicitly, and subtracted
from the analytical force during the simulation. The computational cost of
the self-interaction term is very low comparing with the SPME
force calculation. It is shown that the self-interaction  error
dominate when charge density is low~\cite{cerutti2009staggered}.
Therefore, we always remove the self-interaction term from
the analytical differentiation.
% , and call it as the original analytical
% differetiation when comparing it with the staggered mesh analytical
% differetiation.

@@

% \subsection{The error estimate of the total force}

% In the previous subsections, we estimate the error of direct part and
% reciprocal part separately. Here we provide the error estimate for the
% total force. The mean of the total error force is simply:
% \begin{align}
%   \langle\Delta\v F^{\textrm{x}}(\v r) \rangle
%   &=
%   \langle\Delta\v F_{\textrm{dir}}(\v r) \rangle +
%   \langle\Delta\v F^{\textrm{x}}_{\textrm{rec}}(\v r) \rangle,
% \end{align}
% where \textrm{x}
% denotes any method for the reciprocal Ewald summation, which
% could be truncated Ewald summation (``tr''), ik-/analytical differentiation
% (``ik''/''ana'') or staggered method
% ik-/analytical differentiation (``st,ik''/''st,ana'').
% The mean square total force is
% \begin{align}\nonumber
%   \langle\,\vert\Delta\v F^{\textrm{x}}(\v r)\vert^2\rangle
%   = \,&
%   \langle\,\vert
%   \Delta\v F_{\textrm{dir}}(\v r) + \Delta\v F^{\textrm{x}}_{\textrm{rec}}(\v r) 
%   \vert^2 \rangle \\\nonumber
%   = \,&
%   \langle\,\vert\Delta\v F_{\textrm{dir}}(\v r)\vert^2\rangle +
%   \langle\,\vert\Delta\v F^{\textrm{x}}_{\textrm{rec}}(\v r)\vert^2\rangle +
%   2\,\langle\Delta\v F_{\textrm{dir}}(\v r)\rangle
%   \cdot \langle\Delta\v F^{\textrm{x}}_{\textrm{rec}}(\v r)\rangle \\
%   \label{eqn:error-total}
%   &
%   + 2\,\textrm{Corr}(\Delta\v F_{\textrm{dir}}(\v r),
%   \Delta\v F^{\textrm{x}}_{\textrm{rec}}(\v r))
% \end{align}
% Where $ \textrm{Corr}(\Delta\v F_{\textrm{dir}}(\v r), \Delta\v
% F^{\textrm{x}}_{\textrm{rec}}(\v r))$ denotes the correlation of
% $\Delta\v F_{\textrm{dir}}(\v r)$ and $\Delta\v
% F^{\textrm{x}}_{\textrm{rec}}(\v r)$:
% \begin{align}\label{eqn:error-total-corr}
%   \textrm{Corr}(\Delta\v F_{\textrm{dir}}(\v r),
%   \Delta\v F^{\textrm{x}}_{\textrm{rec}}(\v r))
%   & =
%   \langle
%   \Delta\v F_{\textrm{dir}}(\v r)\cdot
%   \Delta\v F^{\textrm{x}}_{\textrm{rec}}(\v r)
%   \rangle -
%   \langle
%   \Delta\v F_{\textrm{dir}}(\v r)
%   \rangle\cdot
%   \langle
%   \Delta\v F^{\textrm{x}}_{\textrm{rec}}(\v r)
%   \rangle.
% \end{align}
% In Eqn.~\eqref{eqn:error-total}, the estimates of errors
% $\langle\,\vert\Delta\v F_{\textrm{dir}}(\v r)\vert^2\rangle $,
% $\langle\,\vert\Delta\v F^{\textrm{x}}_{\textrm{rec}}(\v
% r)\vert^2\rangle $, $\langle\Delta\v F_{\textrm{dir}}(\v
% r)\rangle$ and $\langle\Delta\v F^{\textrm{x}}_{\textrm{rec}}(\v
% r)\rangle $ have been established.
% In practice, the correlation~\eqref{eqn:error-total-corr}
% is often assumed to be varnished.
% % In section \ref{sec:tmp3}, it is
% % shown that the correlation between the direct and reciprocal error
% % force can be safely neglected.
